---
title: "Class 1 survey analysis"
output: html_document
date: "2024-09-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
C1survey<-read_csv("https://raw.githubusercontent.com/kijohnson/ADA-2024/main/Class%201%20(Getting%20started%20and%20RR%20I)/Lab/Class%201%20Survey%20Fall%202024_di.csv")
```

##
##
2. Determine: a. the number of observations (i.e. the number of people who filled out the survey) and b. the number of variables in the dataframe.    
```{r}
dim(C1survey)
# 29 observations and 27 variables
str(C1survey)
```
##
##
 3. a. Rename the column variables to something shorter and that is descriptive of what the variable is about (for example like_dogs for the 'Do you like dogs?' question variable) and b. Display that they are renamed.
```{r}
names(C1survey)[1:27]<-c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three","public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city", "highest_educ_level")
colnames(C1survey)
```
##
##
 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the C1survey dataset.
```{r}
sapply(C1survey, class) 
table(sapply(C1survey, class))
summary(C1survey)
```
##
##
5. a. Using code, check the bday and bmonth variables for any unusual or missing values. If you find any, b. describe what you will do with the unusual values in a comment before or immediately following the code that does it, and c. after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics
```{r}
# a. code for values for bday and bmonth
C1survey$bday
C1survey$bmonth

# b. We will recode bday with non numeric values

# bday recoding 
class(C1survey$bday)
C1survey$bday<-ifelse(C1survey$bday == "May 21-report 21", "21", C1survey$bday)
C1survey$bday<-ifelse(C1survey$bday == "March 31st", "31", C1survey$bday)
C1survey$bday

# convert to numeric before doing statistics
C1survey$bday<-as.numeric(C1survey$bday)
table(C1survey$bday, useNA = "always") 

# bmonth recoding
class(C1survey$bmonth)

C1survey$bmonth[C1survey$bmonth=="May 21-report 5"]<-"5"
C1survey$bmonth[C1survey$bmonth=="February"]<-"2"
C1survey$bmonth[C1survey$bmonth=="September"]<-"9"
C1survey$bmonth[C1survey$bmonth=="July"]<-"7"
C1survey$bmonth[C1survey$bmonth=="March"]<-"3"
C1survey$bmonth # check values 

# convert to numeric before doing statistics
C1survey$bmonth<-as.numeric(C1survey$bmonth)
C1survey$bmonth # check values 

# c. Next, we will calculate medians for bday and bmonth 
median(C1survey$bday, na.rm=TRUE) # 14 
median(C1survey$bmonth, na.rm=TRUE) # 7
```
##
6. a. Create a new variable called bseason that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). b. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code. 
```{r}
# a.
# we will start by using dplyr
C1survey <- C1survey %>%
  mutate(bseason=case_when(bmonth %in% c(12,1,2) ~"Winter",
                           bmonth %in% c(3,4,5) ~"Spring",
                           bmonth %in% c(6,7,8) ~"Summer",
                           bmonth %in% c(9,10,11) ~"Fall"))
table(C1survey$bmonth, C1survey$bseason)
C1survey$bseason_f<-factor(C1survey$bseason, levels=c("Winter", "Spring", "Summer", "Fall"))

# b. 
table(C1survey$bmonth, C1survey$bseason)
table(C1survey$bmonth, C1survey$bseason_f)

# c.
tab<-addmargins(table(C1survey$bmonth, C1survey$bseason, useNA = "always"), 1)
tab
tab<-addmargins(table(C1survey$bmonth, C1survey$bseason_f, useNA = "always"), 1)
tab

# Births by season: Fall = 4 Winter = 9 Spring = 8 Summer = 8
```
7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #'s).
```{r}
#I would like to see what is the most favorite month among the respondents.To find that i will be calculating the mode (most frequent value) for the variable of interest. Here:The most favorite month
("fav_month" %in% names(C1survey)) 
{
    
    # Create a function to find the mode (most common value)
    find_mode <- function(x) {
      uniq_x <- unique(x)
      uniq_x[which.max(tabulate(match(x, uniq_x)))]
    }
    
    # Find the most common favorite month
    most_common_month <- find_mode(C1survey$fav_month)
    
    # Display the result
    most_common_month
}

```

##